---
title: "Xuan Yu - HW1"
author: "Xuan Yu"
date: "9/3/2018"
output: pdf_document
---


## Problem 1

###This is the summary for the model:
```{r linear model, old}
oldfaithful <- read.csv("/Users/xuanyu/Desktop/MIDS courses/data modeling/HW/HW1/OldFaithful.csv")
lm_old <- lm(Interval~Duration, data = oldfaithful)
summary(lm_old)
```

###This is the 95% confidence interval for the model:
```{r 95% confidence interval, old}
confint(lm_old)
```

###Check the assumption:
We are using the first three plots of the first commond and then the second commond to check the assumption.
```{r residual plot, old}
par(mfcol = c(3,2))
plot(lm_old)
plot(1:dim(oldfaithful)[1], lm_old$residuals) #for the independence assumption
```

###Description:
All the assumptions are met.

###Prediction
Here is the 95% prediction interval when the duration of the previous one is 4 minutes:
```{r 95% prediction interval at 4, old}
newdata.old <- data.frame(Duration = 4)
predict.lm(lm_old, newdata.old, interval = "prediction")
```

###Conclusion:
When the duration of the previous eruption increases by 1 minute, the interval time until the next eruption will increase 10.741 minutes. The following plot shows this relationship.
```{r conclusion, old}
plot(oldfaithful$Duration, oldfaithful$Interval)
abline(33.8282, 10.7410)
```

## Problem 2

###Load in the data and check the assumptions.
```{r linear model, resp}
respiratory <- read.csv("/Users/xuanyu/Desktop/MIDS courses/data modeling/HW/HW1/Respiratory.csv")
```

We found that the linearity and normality assumption are not met.

###Do the transformation and then the linear regression:
```{r linear regression, resp}
respiratory$LogRate <- log(respiratory$Rate)
lm_resp_trans <- lm(LogRate~Age, data = respiratory)
summary(lm_resp_trans)
```

###Check the assumptions again:
```{r assumption after trans, resp}
par(mfcol = c(3,2))
plot(lm_resp_trans)
plot(1:dim(respiratory)[1], lm_resp_trans$residuals) #for the independence assumption
```

###Description:
All the assumptions are met.

###This is the 95% confidence interval for the model:
```{r 95% confidence interval, resp}
confint_log <- confint(lm_resp_trans)
exp(confint_log)
```

###Prediction
Here is the 95% prediction rate for three individual children: a 1 month old, an 18 months old, and a 29 months old:
```{r 95% individual prediction interval at 1, 18, 29, resp}
newdata.resp <- data.frame(Age = c(1, 18, 29))
predict_confint_resp <- predict.lm(lm_resp_trans, newdata.resp, interval = "prediction")
predict_confint_resp <- exp(predict_confint_resp)
cbind(newdata.resp, predict_confint_resp)
```

###Conclusion:
exp(-0.0190090) = 98.117%.

According to the fomula, when the age of the baby increases by 1 month, the respiratory rate will become 98.117% of itself. 

In other word, when the age of the baby increases by 1 month, the respiratory rate will decrease by 1.883%.

The following plot shows this relationship between the respiratory rate and the age:
```{r plot, resp}
respiratory$log_resp <- (-0.0190090 * respiratory$Age) + 3.8451185
plot(respiratory$Age, exp(respiratory$log_resp), type = 'l')
points(respiratory$Age, respiratory$Rate)
```

## Problem 3

###Load in the data and plot it, we see an outlier, we run the model with the outlier, and linearity assumption is not met:
```{r remove PB, ele}
elections <- read.csv("/Users/xuanyu/Desktop/MIDS courses/data modeling/HW/HW1/Elections.csv")
plot(elections$Bush2000, elections$Buchanan2000)
lm_ele <- lm(Bush2000~Buchanan2000, data = elections)
plot(lm_ele)
```

So we decied to remove the outlier to do the prediction:

After checking the linear regression assumptions, we found that the linearity and normality assumption are not met.

###Transformation:
Because the linear regression doesn't met the assumptions, we need to do transformation for it. 

###Here is the transformation and the linear regression for it:
```{r trans, ele}
ele_without_PB <- elections[elections$County != "Palm Beach",]
ele_without_PB$LogY <- log(ele_without_PB$Buchanan2000)
ele_without_PB$LogX <- log(ele_without_PB$Bush2000)
lm_ele_trans <- lm(LogY~LogX, data = ele_without_PB)
```

###Check the assumption again:
```{r check the assumptions after trans, ele}
par(mfcol = c(3,2))
plot(lm_ele_trans)
plot(1:dim(ele_without_PB)[1], lm_ele_trans$residuals) #for the independence assumption
```

###Description:
All assumtions are met.

###This is the summary and confidence interval for the transformed model:
```{r summary and confint, ele}
summary(lm_ele_trans)
confint_log <- confint(lm_ele_trans)
exp(confint_log)
```

###The prediction for Buchanan's votes in Palm Beach County:
```{r prediction, ele}
newdata.ele <- data.frame(LogX = log(152846))
predict_confint_ele <- predict.lm(lm_ele_trans, newdata.ele, interval = "prediction")
predict_confint_ele <- exp(predict_confint_ele)
predict_confint_ele
```

###Votes intended for Gore:
```{r votes for Gore}
c(3407 - predict_confint_ele[3], 3407 - predict_confint_ele[2])
```

###Conclusion:
So Buchanan's votes in Palm Beach county should be in the range of (250.8001, 1399.164), but he got 3407 votes. So there were (2007.836, 3156.200) votes which are intended for Gore.
